Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble

### 2022-09-21 paper/paper.Rmd
close #5 by adding references

Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@REF].
These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.

At the same time, ensemble forecasts have a range of potential downsides.
The most significant is their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
Forecasts from these ensembles can be hard to interpret as it is difficult to link them to the underlying model assumptions.
Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute, and difficulty improving the quality of the ensemble [@REF].
Each of these may impact the long-term quality of the resulting ensemble forecasts.

Here, we focus on addressing the issues of resource cost, and interpretability.
In climate forecasting, emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
This means that non-parametric emulation is less suitable for use in epidemiological forecasting (while noting that epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@REFLSHTMTBEMULATOR]). 
In this work, we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, and limited computational resources with an easily generalised implementation.

@@

The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@REF] in late 2020, and the European forecast hub [@ECDC] in 2021.
All of these collaborations, which the authors (particularly SF and JB) have been involved with managing,  ensembled contributions from multiple independent teams using a similar approach (most recently an unweighted median ensemble [@REF]) and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting [@REF].
However, they have struggled to provide feedback to those contributing forecasts that would allow them to improve their forecast approaches, have found it difficult to improve the quality of their ensemble forecasts [@REF], and have struggled to maintain incentives for forecasters to continue to contribute which may impact the long-term quality of the ensembles they produce.

These forecast collaborations, and other similar infectious disease forecasting efforts, have focussed on forecasting reported cases and deaths by epidemiological week (i.e Saturday to Saturday) over a one to four-week time horizon. Generally, ground truth data is available daily, and is updated in retrospect as new data becomes available. Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].
Most submitted forecasts use available daily data, along with potentially other leading predictors, to produce daily forecasts that are then aggregated into weekly forecasts.
Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts [@REF].
In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF], and human insight based forecast models that may also include elements of other methods [@REF].
Real-time evaluation has shown that each of these classes of models may perform well, although in general, semi-mechanistic models and models that include human insights have outperformed other approaches during certain periods [@REF].
We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods [@REF], and human insight-based forecasts [@REF].

@@

We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@REF].
We used the latest available observed data as of the 20th of August 2022. As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
This follows the practice of the European forecasting hub project in treating the latest available data as the forecast target regardless of the number of potential data revisions [@REF].
Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.

The European forecast hub requests forecasts for one to four-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
No restrictions were placed on who could submit forecasts and the hub team actively invited participation from research groups known to be involved with COVID-19 forecasting projects.
Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent-based models, semi-mechanistic models, expert elicitation models, and ensembles of multiple approaches [@https://covid19forecasthub.eu/community.html].
Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week). 
The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any validly submitted forecast (where validity was defined as passing minimal formatting checks by the hub team and timely submission). 
An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
Submitted forecasts and target observations are available from the European forecast hub GitHub repository [@REF]. We provide code in the repository of this study to streamline access.

@@

We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the European forecast hub  [@REF].
We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).

@@

Our study benefits from having been conducted using forecasts produced in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the European forecast hub).
This means that we can be confident we have captured how our forecast methodology behaves in real-world usage, at least within our study period. 
The downside of this approach is that it was not possible to update the surrogate model over time in response to initial evaluation or to explore other parameterisations that might be more successful of which there are likely several.
However, as our study has been conducted with a focus on reproducibility and openness our findings can be replicated or extended by others regardless of compute availability (due to our use of GitHub actions as a compute platform which is freely available to researchers).
Importantly in this study, we focussed on replicating the forecast hub ensembles' observed behaviour rather than attempting to define an optimal forecast for forecast consumers.
It is possible that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@REF], then our approach makes sense and may be optimal for producing a "good" forecast.
Our focus on replicating the performance of the hub ensemble is also useful as the surrogate model may highlight some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, and the growth rate tending towards zero as the forecast horizon increases.
It also highlights some of the differences between our surrogate forecast model and the ensemble that may lead to new insights in understanding the mechanisms leading to the ensembles behaviour, such as the generally poor coverage of the ensemble that could not be explained by the assumptions we used in developing our surrogate methodology.
Whilst we normalised reported cases to be population-adjusted incidence rates our results are conditional on the use of the weighted interval score as an evaluation metric.
As this proper scoring rule measures absolute performance this means that forecasts during periods of higher incidence are given more weight than forecasts from periods of low incidence.
It also means that overprediction is penalised more than underprediction as incidence rates are bounded at zero but relatively weakly bounded by populations at the upper bound (as incidence rates are typically only a small fraction of the overall population).
This bias could explain the relatively poor performance of the surrogate model, compared to the ensemble, despite the surrogate model being comparably well-calibrated.
We considered alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick relatively closely to the methodology used by the European forecast hub, aside from the use of population weighting in order to facilitate comparison between forecast locations, as these choices inform the development of submitted models and so are key to our findings.

@@

There are no other studies in the epidemiology literature which we are aware of that attempt to develop a forecasting model based on the observed behaviour of a multi-team, multi-model ensemble.
Few studies focus on delivering computationally feasible forecasting models in a reproducible framework backed by an openly accessible compute platform. 
However, the US [@REF], European [@REF], and Germany/Poland [@REF] forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles. 
In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study). 
The poor calibration of the forecast ensembles produced by median Hub ensembles as been noted repeatedly [REFEVAN; REFJOHANNES; REFKATH] but little progress has been made understanding the causes or suggesting alternatives forecasts for policy makers to use.
The US forecast hub identified the top 5 performing models and noted the structural assumptions they made [@REF].
However, they did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model or explore the performance of a forecasting model designed with these assumptions in mind.
Similarly, the Germany and Poland forecasting hubs was able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@REF; @REF; @REF].
In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions after having concluded it was a favourable choice [@REF]. 
The CDC, and European, forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling [@REF].
Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the CDC forecasting hub and did not outperform in the case of the European forecasting hub.
No forecast hub has switched to these alternative ensemble designs for their operational forecast suggesting they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study. 
Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric. 
Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not the model-based forecasts [@refNikosinGermany].

>>>>

Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Sherratt2022-ue].
These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.

At the same time, ensemble forecasts have a range of potential downsides.
The most significant is their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
Forecasts from these ensembles can be hard to interpret as it is difficult to link them to the underlying model assumptions.
Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute [@Bosse2022-hm], and difficulty improving the quality of the ensemble [@Sherratt2022-ue].
Each of these may impact the long-term quality of the resulting ensemble forecasts.

Here, we focus on addressing the issues of resource cost, and interpretability.
In climate forecasting, emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
This means that non-parametric emulation is less suitable for use in epidemiological forecasting (while noting that epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@REFLSHTMTBEMULATOR]).

@@

The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic, partially due to an increased need from stakeholders for reliable forecast models but also due to the perceived low quality of many forecasts otherwise being produced, with the establishment of the CDC forecasting hub [@REF] in early 2020, the Germany/Poland forecast hub [@REF] in late 2020, and the European forecast hub [@Sherratt2022-ue] in 2021.
All of these collaborations, which the authors (particularly SF and JB) have been involved with managing,  ensembled contributions from multiple independent teams using a similar approach (most recently an unweighted median ensemble [@Sherratt2022-ue]) and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
However, they have struggled to provide feedback to those contributing forecasts that would allow them to improve their forecast approaches, have found it difficult to improve the quality of their ensemble forecasts [@Sherratt2022-ue], and have struggled to maintain incentives for forecasters to continue to contribute which may impact the long-term quality of the ensembles they produce.

These forecast collaborations, and other similar infectious disease forecasting efforts, have focussed on forecasting reported cases and deaths by epidemiological week (i.e Saturday to Saturday) over a one to four-week time horizon. Generally, ground truth data is available daily, and is updated in retrospect as new data becomes available. Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].
Most submitted forecasts use available daily data, along with potentially other leading predictors, to produce daily forecasts that are then aggregated into weekly forecasts.
Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts [@REF].
In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well, although in general, semi-mechanistic models and models that include human insights have outperformed other approaches during certain periods [@REF].
We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods [@Bosse2022-hm], and human insight-based forecasts [@Bosse2022-hm].

@@

We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub [@Sherratt2022-ue; @EuroHub] from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@REF].
We used the latest available observed data as of the 20th of August 2022. As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
This follows the practice of the European forecasting hub project in treating the latest available data as the forecast target regardless of the number of potential data revisions [@Sherratt2022-ue].
Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.

The European forecast hub requests forecasts for one to four-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
No restrictions were placed on who could submit forecasts and the hub team actively invited participation from research groups known to be involved with COVID-19 forecasting projects.
Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent-based models, semi-mechanistic models, expert elicitation models, and ensembles of multiple approaches [@https://covid19forecasthub.eu/community.html].
Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week). 
The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any validly submitted forecast (where validity was defined as passing minimal formatting checks by the hub team and timely submission). 
An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
Submitted forecasts and target observations are available from the European forecast hub GitHub repository [@EuroHubGitHub]. We provide code in the repository of this study to streamline access.

@@

We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the European forecast hub  [@Sherratt2022-ue].
We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).

@@

Our study benefits from having been conducted using forecasts produced in real-time, rather than retrospectively, and submitted to an independent forecast research hub (though we note the overlap between authors on this study and the European forecast hub).
This means that we can be confident we have captured how our forecast methodology behaves in real-world usage, at least within our study period. 
The downside of this approach is that it was not possible to update the surrogate model over time in response to initial evaluation or to explore other parameterisations that might be more successful of which there are likely several.
However, as our study has been conducted with a focus on reproducibility and openness our findings can be replicated or extended by others regardless of compute availability (due to our use of GitHub actions as a compute platform which is freely available to researchers).
Importantly in this study, we focussed on replicating the forecast hub ensembles' observed behaviour rather than attempting to define an optimal forecast for forecast consumers.
It is possible that if we had instead aimed to develop a forecast methodology that minimised the evaluation criteria we planned to use, especially if we relaxed our assumed compute resource constraints, we would have produced forecasts that performed better relative to the hub ensemble.
However, if we start from the view that the forecast hub ensemble has traits that are desirable for use by policy-makers (i.e robustness and good average performance), which can be found widely in the literature [@REF; [@Sherratt2022-ue]], then our approach makes sense and may be optimal for producing a "good" forecast.
Our focus on replicating the performance of the hub ensemble is also useful as the surrogate model may highlight some of the emergent behaviour of the ensemble captured in our assumptions, such as auto-correlation across time points, and the growth rate tending towards zero as the forecast horizon increases.
It also highlights some of the differences between our surrogate forecast model and the ensemble that may lead to new insights in understanding the mechanisms leading to the ensembles behaviour, such as the generally poor coverage of the ensemble that could not be explained by the assumptions we used in developing our surrogate methodology.
Whilst we normalised reported cases to be population-adjusted incidence rates our results are conditional on the use of the weighted interval score as an evaluation metric.
As this proper scoring rule measures absolute performance this means that forecasts during periods of higher incidence are given more weight than forecasts from periods of low incidence.
It also means that overprediction is penalised more than underprediction as incidence rates are bounded at zero but relatively weakly bounded by populations at the upper bound (as incidence rates are typically only a small fraction of the overall population).
This bias could explain the relatively poor performance of the surrogate model, compared to the ensemble, despite the surrogate model being comparably well-calibrated.
We considered alternative methods of forecast evaluation that would be robust to this potential source of bias but choose to stick relatively closely to the methodology used by the European forecast hub [@Sherratt2022-ue], aside from the use of population weighting in order to facilitate comparison between forecast locations, as these choices inform the development of submitted models and so are key to our findings.

@@

There are no other studies in the epidemiology literature which we are aware of that attempt to develop a forecasting model based on the observed behaviour of a multi-team, multi-model ensemble.
Few studies focus on delivering computationally feasible forecasting models in a reproducible framework backed by an openly accessible compute platform. 
However, the US [@REF], European [@Sherratt2022-ue], and Germany/Poland [@REF] forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles. 
In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study). 
The poor calibration of the forecast ensembles produced by median Hub ensembles as been noted repeatedly [REFEVAN; REFJOHANNES; @Sherratt2022-ue] but little progress has been made understanding the causes or suggesting alternatives forecasts for policy makers to use.
The US forecast hub identified the top 5 performing models and noted the structural assumptions they made [@REF].
However, they did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model or explore the performance of a forecasting model designed with these assumptions in mind.
Similarly, the Germany and Poland forecasting hubs was able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@REF; @REF; @Sherratt2022-ue].
In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions. 
The CDC, and European, forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling [@REF; @Sherratt2022-ue].
Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases in the case of the CDC forecasting hub and did not outperform in the case of the European forecasting hub.
No forecast hub has switched to these alternative ensemble designs for their operational forecast suggesting they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study. 
Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric. 
Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not the model-based forecasts [@Bosse2022-hm].
